---
title: "Homework 2 Report"
output:
  html_document: default
  pdataset_document: default
---


## Name

Vijetha Ramdas

### Setup

Read in the data with `read_csv()` and store the data as an R object named `dataset`. Check the data to make sure all of the expected observations and variables are there.


```{r}
## Load the data and any necessary packages here.
library(tidyverse)
dataset <- read_csv("../maacs.csv.gz")
head(dataset)
tail(dataset)
nrow(dataset)
```

## Part 1

We will first consider the relationship between FEV1 and age. In general, it is expected that as children get older (and hence, larger in size), their FEV1 values should get higher. 

Consider the statement "FEV1 values in children are higher in older children relative to younger children". 

Write a function in R that takes the `dataset` object as an argument and returns `TRUE` if the statement above is true for the dataset and `FALSE` otherwise. 

NOTE: In order to write this function, you will need to translate the statement above into something that can be checked with the data. There are many ways in which you can do that translation correctly and you only need to pick one way here. 

NOTE: For this part, do not use any plots.

```{r}
## Write your function here

test_statement <- function(dataset) {
  # Find the median age to split into the old and young groups
  mid <- dataset$age |> median() 
  dataset = dataset |>
    mutate(group = ifelse(age >= mid, "Old", "Young"))
  
  # Calculate the mean for each group 
  mean_table <- dataset |> 
    group_by(group) |>
    summarise(mean(fev1, na.rm = TRUE))
  
  # Calculate the difference between groups
  mean_diff <- mean_table[1,2] - mean_table[2,2]

  # Return True if the Old group has a higher mean by at least .5
  if (mean_diff > .5) {
    return(TRUE)}
  else{
    return(FALSE)}
}

test_statement(dataset)

```


## Part 2

Fit a linear regression model with FEV1 as the outcome and age as a predictor.

How much does FEV1 change for a 1-year increase in the child's age? 


```{r}
## Add your code here

model <- lm(fev1 ~ age, data = dataset)
summary(model)  # view the model coefficient/results
```


Write your data analysis statement interpreting the regression model here:

For a one year increase in the child's age, FEV1 increases by 0.171162 L.

## Part 3

Develop **three** supporting premises derived from the data that support the statement you wrote in Part 2. These can be plots, other summary statistics, or model results.

NOTE: At least one supporting premise should use a plot.

```{r}
## Add your code here

# find the median age to define the Old and Young group
mid <- dataset$age |> median()
print(mid)

dataset = dataset |>
  mutate(group = ifelse(age >= mid, "Old", "Young"))
  
# Premise 1: scatterplot between age and fev1
ggplot(dataset) +
  geom_point(aes(x = age, y = fev1)) +
  labs(x = "Age", y = "FEV1 (L)")

# Premise 2: boxplot of fev1 distribution split by age group
ggplot(dataset) +
  geom_boxplot(aes(x = fev1, y = group)) +
  labs(x = "FEV1 (L)", y = "Age Group")

# Premise 3: histogram of fev1 distribution split by age group
ggplot(dataset) +
  geom_histogram(aes(x = fev1), bins = 15) +
  facet_wrap(~ group, nrow = 2) +
  labs(x = "FEV1 (L)", y = "Count")
```



Write the three supporting premise statements here:

1. On the scatterplot, children with higher ages tend to also have a higher FEV1 value. There is a clear positive association between the age and FEV1 variables on the plot.

2. The median FEV1 value of the younger age group box plot is lower than the median FEV1 value of the older age group. The boxplots are also relatively symmetric, and the entire range of FEV1 values for the younger group is lower than the median FEV1 value of the older age group.

3. The center of the histogram showing the FEV1 distribution of the younger group is lower than the center of the histogram showing the FEV1 distribution of the older group. The histograms are relatively symmetric, and only the older group has very high ouliers.

## Part 4

For each of the supporting premises above, write a function that takes the `dataset` object as an argument and returns `TRUE` if the supporting premise statement above is true for the dataset and `FALSE` otherwise. 

For statements involving plots, instead of returning `TRUE` or `FALSE`, you function should do two things:

1. Produce the plot that is used in the statement

2. Produce a hypothetical version of the plot in the event that the statement is true. This can be done using simulated data or by simply hand drawing a plot.


```{r}
set.seed(19)
## Function for supporting premise statement 1
premise_1 <- function(dataset){
  # produce the original plot
  real_plot <- ggplot(dataset) +
    geom_point(aes(x = age, y = fev1)) +
    labs(x = "Age", y = "FEV1 (L)")

  # generate distribution under the statement (positive relationship bewteen fev1 and age)
  x = runif(150, min(dataset$age), max(dataset$age))
  y = .25*x + rnorm(75, 0, .5) 
  fake_dist <- tibble(x, y)
  
  # produce the plot with the simulated data
  fake_plot <- ggplot(fake_dist) +
    geom_point(aes(x = x, y = y)) +
    labs(x = "Age", y = "FEV1 (L)")
  
  print(real_plot)
  print(fake_plot)
}

premise_1(dataset)
```


```{r}
## Function for supporting premise statement 2
premise_2 <- function(dataset){
  # produce the original plot
  real_plot <- ggplot(dataset) +
  geom_boxplot(aes(x = fev1, y = group)) +
  labs(x = "FEV1 (L)", y = "Age Group")
  
  # create the simulated data where Younger childer have a lower mean fev1 level than older children
  fake_data <- tibble(
    x = c(rnorm(75, 1.3, .35), rnorm(75, 2.5, .35)),
    y = c(rep("Young", 75), rep("Old", 75))
  )
  
  # produce the plot with the simulated data
  fake_plot <- ggplot(fake_data) +
  geom_boxplot(aes(x = x, y = y)) +
  labs(x = "FEV1 (L)", y = "Age Group")
  
  print(real_plot)
  print(fake_plot)
}

premise_2(dataset)
```


```{r}
## Function for supporting premise statement 3
premise_3 <- function(dataset){
  # produce the original plot 
  real_plot <- ggplot(dataset) +
  geom_histogram(aes(x = fev1), bins = 15) +
  facet_wrap(~ group, nrow = 2) +
  labs(x = 'Count', y = "FEV1 (L)")
  
  # create the simulated data where Younger childer have a lower mean fev1 level than older children
  fake_data <- tibble(
    x = c(rnorm(75, 1.3, .35), rnorm(75, 2.5, .35)),
    y = c(rep("Young", 75), rep("Old", 75))
  )
  
  # produce the plot with the simulated data
  fake_plot <- ggplot(fake_data) +
  geom_histogram(aes(x = x), bins = 15) +
  facet_wrap(~ y, nrow = 2) +
  labs(x = 'Count', y = "FEV1 (L)")
  
  print(real_plot)
  print(fake_plot)
}

premise_3(dataset)
```

Execute each of your function and show that the produce the expected output.

## Part 5

Describe one alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children". 

Create a fault tree for the alternative outcome describing how the alternative outcome could be realized in the data even if the primary statement were true.

Your fault tree should be created as a separate image and does not need to be created in R. Upload the image of the fault tree to Canvas.

Alternative: FEV1 values in children are the same in older children and younger children.
